---
# This file generates AGENTS.md in the package root.
# It is not part of the package build or the pkgdown site.
# To regenerate it, run from the package root:
#     rmarkdown::render("vignettes/_AGENTS.Rmd", output_file = file.path("AGENTS.md"), output_dir = ".")
output:
  github_document:
    html_preview: false
params:
  url: "../SHARED_AGENTS.md"
---
# teal R Package Development Guide

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Package Overview

`teal` is the main package of the framework to provide shiny web apps for analyzing clinical trials data. 
It provides a unified user interface and features for the different modules that analyze the data.

The most important functions are:

- `init`: Creates a teal application.
- `modules`: Combines analyses implemented as Shiny or teal modules.
- `teal_slices`: Filters datasets.

Other important helpers:
- `example_module`: To show and test framework features.
- `modify_*` and `disable_*` functions to alter the UI of the app and remove some features respectively. 
- `*_decorators`:  Modules can be extended with decorators this functions make it easier to build custom modules using it.
- `validate*`: Functions to validate user's input or objects used in modules.

## Development Context

The main feature of this package is to generate `teal` apps and integrate their features for the users.


### Direct dependencies

- **shiny** - Web app generation and isolating framework. 
- **teal.code** - Bare code generation and evaluation ensuring reproducibility
- **teal.data** - Data management and relationships between datasets via join_keys
(contains sample data for ADaM datasets and default keys to merge ADaM datasets)
- **teal.reporter** - Report generation functionality
- **teal.slice** - Data filtering capabilities for application
- **teal.widgets** - Reusable UI components for user input and output
- **teal.logger** - Standardized logging across the framework

### Supporting Packages

The `teal` package is used by module packages in examples and functions:

- **teal.modules.general** (tmg) - General-purpose analysis modules
- **teal.modules.clinical** (tmc) - Clinical trial specific modules
- **teal.modules.hermes** - analysis modules using MultiAssayExperiment from Bioconductor.
- **teal.goshawk** - Pharmacokinetics analysis modules
- **teal.osprey** - Advanced clinical analysis modules

In addition to help the analysis these two packages might be used on modules:

- **teal.picks** - Data selection and merging utilities using `teal.data` objects
- **teal.transform** - Data transformation utilities (deprecated in favor of teal.picks)


## Workflows

- Before fixing an issue consider that this package is part of teal framework, so a bug that manifests in this package might have a different origin.
- Before changing code in the `init()` function, consider the compatibility of this change with the module packages.
- The internal `TealAppDriver` object is used for end to end test the modules packages. If any change is done check the compatibility with the module packages.
- If something is needed for more than one module it might be needed on this package to be exported to all of them or in one of the dependencies.
- The package should be back-compatible
- Use the pattern `module_<name>.R` for shiny modules.
- Prefix helper functions with the main function they support
- Group related utilities in shared files (e.g., `utils.R`, `validations.R`).

```{r teal_content, echo=FALSE, results='asis'}
# Append common shared instructions from teal framework (special case as they are on this repo)
common_claude_shared <- readLines(params$url)
cat(common_claude_shared, sep = "\n")
```
