---
title: "Adding support for Reporting to custom modules"
author: "NEST CoreDev"
date: "2022-05-23"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Adding support for Reporting to custom modules}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

`teal` supports an in-built reporting feature using the `vignette("teal.reporter")` package. Head to its documentation
if you want to know more about the reporting itself.

This article is targeted to module developers and explains how to enhance a custom `teal` module with an automatic
reporting feature. The enhancement allows users to add snapshots of the module outputs to a report and review it in
another module that is automatically provided by `teal` and designed to let users interact with the report.

The responsibilities of a module developer include:

- adding the support for reporting to their module
- specifying outputs that go into a snapshot of their module.

The life cycle of objects involved in creation of the report and setting up the module to preview the report is
handled by `teal`.

## Custom module

Let us consider the example module from `teal`:
```{r, message=FALSE}
library(teal)
teal_example_module <- function(label = "example teal module") {
  checkmate::assert_string(label)
  module(
    label,
    server = function(id, data) {
      checkmate::assert_class(data, "tdata")
      moduleServer(id, function(input, output, session) {
        output$text <- renderPrint(data[[input$dataname]]())
      })
    },
    ui = function(id, data) {
      ns <- NS(id)
      teal.widgets::standard_layout(
        output = verbatimTextOutput(ns("text")),
        encoding = selectInput(ns("dataname"), "Choose a dataset", choices = names(data))
      )
    },
    datanames = "all"
  )
}
```

`teal` can launch this example module with the following lines:
```{r, eval = FALSE}
app <- init(
  data = teal_data(
    dataset("IRIS", iris),
    dataset("MTCARS", mtcars)
  ),
  modules = teal_example_module()
)

if (interactive()) shinyApp(app$ui, app$server)
```

## Add support for Reporting

### Change the declaration of the server function

The first step is to add another argument to the server function declaration - `reporter`. This will tell `teal`
that `reporter` is needed by this module and it will be included when module is called. See below:

```{r}
example_module_with_reporting <- function(label = "example teal module") {
  checkmate::assert_string(label)
  module(
    label,
    server = function(id, data, reporter) {
      checkmate::assert_class(data, "tdata")
      moduleServer(id, function(input, output, session) {
        output$text <- renderPrint(data[[input$dataname]]())
      })
    },
    ui = function(id, data) {
      ns <- NS(id)
      teal.widgets::standard_layout(
        output = verbatimTextOutput(ns("text")),
        encoding = selectInput(ns("dataname"), "Choose a dataset", choices = names(data))
      )
    },
    datanames = "all"
  )
}
```

Such a module is ready to be launched again by `teal`:

```{r}
app <- init(
  data = teal_data(
    dataset("IRIS", iris),
    dataset("MTCARS", mtcars)
  ),
  modules = example_module_with_reporting()
)

if (interactive()) shinyApp(app$ui, app$server)
```

`teal` added another tab to the application titled `Report previewer` but besides that there appears to be no change in
how the module works and what it looks like. Users cannot interact with it to add to the report from this module yet.
Thankfully, `teal.reporter` provides `ui` and `server` objects that support that.

### Introduce the new `UI` and the supporting `shiny` modules

We will use `teal.reporter::simple_reporter_ui` and `teal.reporter::simple_reporter_srv` to set up the `UI` and
the `shiny` module that allow users to add content from `example_module_with_reporting` to the report.

```{r}
example_module_with_reporting <- function(label = "example teal module") {
  checkmate::assert_string(label)
  module(
    label,
    server = function(id, data, reporter) {
      checkmate::assert_class(data, "tdata")
      moduleServer(id, function(input, output, session) {
        teal.reporter::simple_reporter_srv(
          id = "reporter",
          reporter = reporter,
          card_fun = function(card) card
        )
        output$text <- renderPrint(data[[input$dataname]]())
      })
    },
    ui = function(id, data) {
      ns <- NS(id)
      teal.widgets::standard_layout(
        output = tagList(
          teal.reporter::simple_reporter_ui(ns("reporter")),
          verbatimTextOutput(ns("text"))
        ),
        encoding = selectInput(ns("dataname"), "Choose a dataset", choices = names(data))
      )
    },
    datanames = "all"
  )
}
```

This module is ready to be launched:
```{r}
app <- init(
  data = teal_data(
    dataset("IRIS", iris),
    dataset("MTCARS", mtcars)
  ),
  modules = example_module_with_reporting()
)

if (interactive()) shinyApp(app$ui, app$server)
```

The new `UI` is visible and the buttons are clickable. An application user can review the card in the `Report previewer`
module and it will appear empty because, as a module developer, we have not yet added any content to the card
of our module.

### Add content to the card

We will use the public API exposed by the `TealReportCard` class to add content to a card. The `teal.reporter::simple_reporter_srv` module accepts the `card_fun` argument that
dictates the way the output from our custom module will look. `ReportCard` and its derivatives add
the content sequentially according to the order of the calls to its methods. The content itself can be explored
by calling the `$get_content` method. If you want to learn more, check out `TealReportCard`'s
documentation and `teal.reporter::ReportCard`.

We will add simple text to the card by modifying the `card_fun` argument passed to `teal.reporter::simple_reporter_srv`.
Make sure to return the `card` object from the passed function, otherwise `teal` might encounter errors.

```{r}
custom_function <- function(card = teal.reporter::ReportCard$new()) {
  card$append_text("This is content from a custom teal module!")
  card
}

example_module_with_reporting <- function(label = "example teal module") {
  checkmate::assert_string(label)
  module(
    label,
    server = function(id, data, reporter) {
      checkmate::assert_class(data, "tdata")
      moduleServer(id, function(input, output, session) {
        teal.reporter::simple_reporter_srv(id = "simpleReporter", reporter = reporter, card_fun = custom_function)
        output$text <- renderPrint(data[[input$dataname]]())
      })
    },
    ui = function(id, data) {
      ns <- NS(id)
      teal.widgets::standard_layout(
        output = tagList(
          teal.reporter::simple_reporter_ui(ns("simpleReporter")),
          verbatimTextOutput(ns("text"))
        ),
        encoding = selectInput(ns("dataname"), "Choose a dataset", choices = names(data))
      )
    },
    datanames = "all"
  )
}
```

```{r}
app <- init(
  data = teal_data(
    dataset("IRIS", iris),
    dataset("MTCARS", mtcars)
  ),
  modules = example_module_with_reporting()
)

if (interactive()) shinyApp(app$ui, app$server)
```

Now, an application user can see the text added by `custom_function` in the `Report previewer` module.

### Add non-text content to the card

`teal.reporter` supports adding tables, charts and more. Explore the API of `teal.reporter::ReportCard` to learn what
types of content are supported.

### `TealReportCard`

`teal` exports `TealReportCard` which is an extension of `teal.reporter::ReportCard` class. `TealReportCard` adds a number of convenience methods to make it work with `teal` features like the filter panel or source code easier. Check out its documentation to learn more at `TealReportCard`.

To support `TealReportCard`, the function that is passed to `teal.reporter::simple_reporter_srv` needs to define
a default value for the card, like below:

```{r}
custom_fun <- function(card = TealReportCard$new()) {
  card
}
```

Otherwise, the API of `TealReportCard` will not be available inside the function.

## Example

Summing up, we could build a regular teal app with code reproducibility and reporter functionality.
Note that the `server` function requires the `filter_panel_api` argument so that the filter panel state can be added to the report.

```{r}
library(teal)
library(teal.reporter)

example_reporter_module <- function(label = "Example") {
  module(
    label,
    server = function(id, data, reporter, filter_panel_api) {
      with_filter <- !missing(filter_panel_api) && inherits(filter_panel_api, "FilterPanelApi")
      checkmate::assert_class(data, "tdata")
      moduleServer(id, function(input, output, session) {
        dat <- reactive(data[[input$dataname]]())
        output$nrow_ui <- renderUI({
          sliderInput(session$ns("nrow"), "Number of rows:", 1, nrow(data[[input$dataname]]()), 10)
        })

        table_q <- reactive({
          req(input$nrow)

          teal.code::new_qenv(tdata2env(data), code = get_code(data)) %>%
            teal.code::eval_code(
              substitute(
                result <- head(data, nrows),
                list(
                  data = as.name(input$dataname),
                  nrows = input$nrow
                )
              )
            )
        })

        output$table <- renderTable(table_q()[["result"]])

        ### REPORTER
        card_fun <- function(card = ReportCard$new(), comment) {
          card$set_name("Table Module")
          card$append_text(paste("Selected dataset", input$dataname), "header2")
          card$append_text("Selected Filters", "header3")
          if (with_filter) {
            card$append_text(filter_panel_api$get_filter_state(), "verbatim")
          }
          card$append_text("Encoding", "header3")
          card$append_text(
            yaml::as.yaml(
              stats::setNames(lapply(c("dataname", "nrow"), function(x) input[[x]]), c("dataname", "nrow"))
            ),
            "verbatim"
          )
          card$append_text("Module Table", "header3")
          card$append_table(table_q()[["result"]])
          card$append_text("Show R Code", "header3")
          card$append_text(paste(teal.code::get_code(table_q()), collapse = "\n"), "verbatim")
          if (!comment == "") {
            card$append_text("Comment", "header3")
            card$append_text(comment)
          }
          card
        }
        teal.reporter::add_card_button_srv("addReportCard", reporter = reporter, card_fun = card_fun)
        teal.reporter::download_report_button_srv("downloadButton", reporter = reporter)
        teal.reporter::reset_report_button_srv("resetButton", reporter)
        ###
      })
    },
    ui = function(id, data) {
      ns <- NS(id)
      teal.widgets::standard_layout(
        output = tableOutput(ns("table")),
        encoding = tagList(
          div(
            teal.reporter::add_card_button_ui(ns("addReportCard")),
            teal.reporter::download_report_button_ui(ns("downloadButton")),
            teal.reporter::reset_report_button_ui(ns("resetButton"))
          ),
          selectInput(ns("dataname"), "Choose a dataset", choices = names(data)),
          uiOutput(ns("nrow_ui"))
        )
      )
    },
    datanames = "all"
  )
}

app <- init(
  data = teal_data(
    dataset("AIR", airquality, code = "data(airquality); AIR <- airquality"),
    dataset("IRIS", iris, code = "data(iris); IRIS <- iris"),
    check = FALSE
  ),
  modules = list(
    example_reporter_module(label = "with Reporter"),
    example_module(label = "without Reporter")
  ),
  filter = list(AIR = list(Month = c(5, 5))),
  header = "Example teal app with reporter"
)

if (interactive()) shinyApp(app$ui, app$server)
```
