---
title: "Adding support for Reporting to custom modules"
author: "NEST CoreDev"
date: "2022-05-23"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Adding support for Reporting to custom modules}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

The `teal` package offers an integrated reporting feature utilizing the `teal.reporter` package. For a comprehensive explanation of the reporting functionality itself, please refer to the documentation therein.

This article is intended for module developers and aims to provide guidance on enriching a custom `teal` module with an automatic reporting feature. This enhancement enables users to incorporate snapshots of the module outputs into a report, which can then be reviewed in another module automatically provided by `teal`. This interaction allows users to seamlessly interact with the report.

The responsibilities of a module developer include:

- Adding support for reporting to their module.
- Specifying the outputs that contribute to a snapshot of their module.

The entire lifecycle of objects involved in creating the report and configuring the module to preview the report is handled by `teal`.

## Custom module

Let us consider the example module from `teal`:
```{r, message=FALSE}
library(teal)
teal_example_module <- function(label = "example teal module") {
  checkmate::assert_string(label)
  module(
    label,
    server = function(id, data) {
      checkmate::assert_class(data, "tdata")
      moduleServer(id, function(input, output, session) {
        output$text <- renderPrint(data[[input$dataname]]())
      })
    },
    ui = function(id, data) {
      ns <- NS(id)
      teal.widgets::standard_layout(
        output = verbatimTextOutput(ns("text")),
        encoding = selectInput(ns("dataname"), "Choose a dataset", choices = names(data))
      )
    },
    filters = "all"
  )
}
```

Using `teal`, you can launch this example module with the following:

```{r, eval = FALSE}
app <- init(
  data = teal_data(
    dataset("IRIS", iris),
    dataset("MTCARS", mtcars)
  ),
  modules = teal_example_module()
)

if (interactive()) shinyApp(app$ui, app$server)
```

## Add support for Reporting

### Modify the declaration of the server function

The first step is to add an additional argument to the server function declaration - `reporter`. This informs `teal` that the module requires `reporter`, and it will be included when the module is called. See below:

```{r}
example_module_with_reporting <- function(label = "example teal module") {
  checkmate::assert_string(label)
  module(
    label,
    server = function(id, data, reporter) {
      checkmate::assert_class(data, "tdata")
      moduleServer(id, function(input, output, session) {
        output$text <- renderPrint(data[[input$dataname]]())
      })
    },
    ui = function(id, data) {
      ns <- NS(id)
      teal.widgets::standard_layout(
        output = verbatimTextOutput(ns("text")),
        encoding = selectInput(ns("dataname"), "Choose a dataset", choices = names(data))
      )
    },
    filters = "all"
  )
}
```

With these modifications, the module is now ready to be launched with `teal`:

```{r}
app <- init(
  data = teal_data(
    dataset("IRIS", iris),
    dataset("MTCARS", mtcars)
  ),
  modules = example_module_with_reporting()
)

if (interactive()) shinyApp(app$ui, app$server)
```

`teal` adds another tab to the application, titled `Report previewer`. However, there is no visible change in how the module operates and appears and the user cannot add content to the report from this module. That requires inserting `teal.reporter` `ui` and `server` elements into the module body.

### Insert `UI` and supporting `shiny` modules for adding report cards

To enable users to add content from `example_module_with_reporting` to the report, we will utilize `teal.reporter::simple_reporter_ui` and `teal.reporter::simple_reporter_srv` to set up the `UI` and the `shiny` module.

```{r}
example_module_with_reporting <- function(label = "example teal module") {
  checkmate::assert_string(label)
  module(
    label,
    server = function(id, data, reporter) {
      checkmate::assert_class(data, "tdata")
      moduleServer(id, function(input, output, session) {
        teal.reporter::simple_reporter_srv(
          id = "reporter",
          reporter = reporter,
          card_fun = function(card) card
        )
        output$text <- renderPrint(data[[input$dataname]]())
      })
    },
    ui = function(id, data) {
      ns <- NS(id)
      teal.widgets::standard_layout(
        output = tagList(
          teal.reporter::simple_reporter_ui(ns("reporter")),
          verbatimTextOutput(ns("text"))
        ),
        encoding = selectInput(ns("dataname"), "Choose a dataset", choices = names(data))
      )
    },
    filters = "all"
  )
}
```

This updated module is now ready to be launched:

```{r}
app <- init(
  data = teal_data(
    dataset("IRIS", iris),
    dataset("MTCARS", mtcars)
  ),
  modules = example_module_with_reporting()
)

if (interactive()) shinyApp(app$ui, app$server)
```

The new `UI` is visible, and the buttons are clickable. An application user can review the card in the `Report previewer` module, but it will appear empty since we haven't added any content to the card of our module yet.

### Add content to the card

To add content to a card, we will utilize the public API exposed by the `TealReportCard` class. The `teal.reporter::simple_reporter_srv` module accepts the card_fun argument that determines the appearance of the output from our custom module. `ReportCard` and its derivatives allow the sequential addition of content according to the order of method calls. To explore the content, we can use the `$get_content` method. For further details, refer to the documentation of `TealReportCard` and `teal.reporter::ReportCard`.

We will add simple text to the card by modifying the `card_fun` argument passed to `teal.reporter::simple_reporter_srv`. 
Remember to return the `card` object from the function, as failure to do so may cause errors in `teal`.

```{r}
custom_function <- function(card = teal.reporter::ReportCard$new()) {
  card$append_text("This is content from a custom teal module!")
  card
}

example_module_with_reporting <- function(label = "example teal module") {
  checkmate::assert_string(label)
  module(
    label,
    server = function(id, data, reporter) {
      checkmate::assert_class(data, "tdata")
      moduleServer(id, function(input, output, session) {
        teal.reporter::simple_reporter_srv(id = "simpleReporter", reporter = reporter, card_fun = custom_function)
        output$text <- renderPrint(data[[input$dataname]]())
      })
    },
    ui = function(id, data) {
      ns <- NS(id)
      teal.widgets::standard_layout(
        output = tagList(
          teal.reporter::simple_reporter_ui(ns("simpleReporter")),
          verbatimTextOutput(ns("text"))
        ),
        encoding = selectInput(ns("dataname"), "Choose a dataset", choices = names(data))
      )
    },
    filters = "all"
  )
}
```

```{r}
app <- init(
  data = teal_data(
    dataset("IRIS", iris),
    dataset("MTCARS", mtcars)
  ),
  modules = example_module_with_reporting()
)

if (interactive()) shinyApp(app$ui, app$server)
```

Now, an application user can see the text added by `custom_function` in the `Report previewer` module.

### Add non-text content to the card

`teal.reporter` supports the addition of tables, charts, and more. For more information, explore the API of `teal.reporter::ReportCard` to learn about the supported content types.

### `TealReportCard`

`teal` exports `TealReportCard`, which extends the `teal.reporter::ReportCard` class and provides several convenient methods to facilitate working with `teal` features like the filter panel or source code. For more details, refer to the documentation of `TealReportCard`.

To support `TealReportCard`, the function that is passed to `teal.reporter::simple_reporter_srv`  must define a default value for the card, as shown below:

```{r}
custom_fun <- function(card = TealReportCard$new()) {
  card
}
```

Without this definition, the API of `TealReportCard` will not be available within the function.

## Example

In conclusion, we have demonstrated how to build a standard `teal` app with code reproducibility and reporter functionality. Note that the `server` function requires the `filter_panel_api` argument so that the filter panel state can be added to the report.

In the final example, we have incorporated `teal.code` snippets. `teal.code` is an R library that offers utilities for storing code and associating it with an execution environment. Within the `ReporterCard`, you will have access to the code necessary to generate the table, as well as the ability to view the content of the generated table. To learn more about `teal.code` see `vignette("qenv", package = "teal.code")`,

```{r}
library(teal)
library(teal.reporter)

example_reporter_module <- function(label = "Example") {
  module(
    label,
    server = function(id, data, reporter, filter_panel_api) {
      with_filter <- !missing(filter_panel_api) && inherits(filter_panel_api, "FilterPanelApi")
      checkmate::assert_class(data, "tdata")
      moduleServer(id, function(input, output, session) {
        dat <- reactive(data[[input$dataname]]())
        output$nrow_ui <- renderUI({
          sliderInput(session$ns("nrow"), "Number of rows:", 1, nrow(data[[input$dataname]]()), 10)
        })

        table_q <- reactive({
          req(input$nrow)

          teal.code::new_qenv(tdata2env(data), code = get_code(data)) %>%
            teal.code::eval_code(
              substitute(
                result <- head(data, nrows),
                list(
                  data = as.name(input$dataname),
                  nrows = input$nrow
                )
              )
            )
        })

        output$table <- renderTable(table_q()[["result"]])

        ### REPORTER
        card_fun <- function(card = ReportCard$new(), comment) {
          card$set_name("Table Module")
          card$append_text(paste("Selected dataset", input$dataname), "header2")
          card$append_text("Selected Filters", "header3")
          if (with_filter) {
            card$append_text(filter_panel_api$get_filter_state(), "verbatim")
          }
          card$append_text("Encoding", "header3")
          card$append_text(
            yaml::as.yaml(
              stats::setNames(lapply(c("dataname", "nrow"), function(x) input[[x]]), c("dataname", "nrow"))
            ),
            "verbatim"
          )
          card$append_text("Module Table", "header3")
          card$append_table(table_q()[["result"]])
          card$append_text("Show R Code", "header3")
          card$append_text(paste(teal.code::get_code(table_q()), collapse = "\n"), "verbatim")
          if (!comment == "") {
            card$append_text("Comment", "header3")
            card$append_text(comment)
          }
          card
        }
        teal.reporter::add_card_button_srv("addReportCard", reporter = reporter, card_fun = card_fun)
        teal.reporter::download_report_button_srv("downloadButton", reporter = reporter)
        teal.reporter::reset_report_button_srv("resetButton", reporter)
        ###
      })
    },
    ui = function(id, data) {
      ns <- NS(id)
      teal.widgets::standard_layout(
        output = tableOutput(ns("table")),
        encoding = tagList(
          div(
            teal.reporter::add_card_button_ui(ns("addReportCard")),
            teal.reporter::download_report_button_ui(ns("downloadButton")),
            teal.reporter::reset_report_button_ui(ns("resetButton"))
          ),
          selectInput(ns("dataname"), "Choose a dataset", choices = names(data)),
          uiOutput(ns("nrow_ui"))
        )
      )
    },
    filters = "all"
  )
}

app <- init(
  data = teal_data(
    dataset("AIR", airquality, code = "data(airquality); AIR <- airquality"),
    dataset("IRIS", iris, code = "data(iris); IRIS <- iris"),
    check = FALSE
  ),
  modules = list(
    example_reporter_module(label = "with Reporter"),
    example_module(label = "without Reporter")
  ),
  filter = list(AIR = list(Month = c(5, 5))),
  header = "Example teal app with reporter"
)

if (interactive()) shinyApp(app$ui, app$server)
```
