---
title: "Customizing Module Output"
author: "NEST CoreDev"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Customizing Module Output}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(teal)
library(ggplot2)
```


## Introduction

The outputs produced by `teal` modules, like graphs or tables, are created by the module developer and look a certain way.
It is impossible to design an output that will satisfy every possible user, so the form of the output should be considered a default value that can be customized.
Here we describe the concept of _decoration_, whereby a `teal_transform_module` can modify an output created by a `teal_module`, enabling you to tailor outputs to your specific requirements without rewriting the original module code.

While it is not required, reading [this vignette](data-transform-as-shiny-module.html) may give you a fuller understanding.


## Decorators

Decorators are created with `teal_transform_module` and thus they are `shiny` modules.
They are passed to `teal_module` constructors as arguments (see below).
Their server logic will be used to modify objects such as plots or tables that exist in the server function of a `teal_module`.
A `ui` function can provide interactivity but that is optional, an app developer is free to use decorator modules that do not require user input.


### Requirements and Limitations

Using decorators requires the following:

1. **Module Support**:<br>
`teal` will apply decorators to `teal_module` outputs but the module in question must explicitly support this functionality.
It the responsibility of to the module developer to ensure that decorators can be used.
2. **Matching Object Names**:<br>
Decorators will reference variables that exist in the `teal_module` server function and therefore must use the appropriate variable names.
Module developers are encouraged to provide the relevant names in the module's documentation, otherwise the person witing a decorator must follow the source code.
3. **Maintaining Object Classes**:<br>
A decorator must not alter the class of the object that it modifies.
This is because a different class may require a different rendering function and that is part of the module structure, which beyond the control of decorators.


#### Example module that we will decorate

```{r undecorated_app}
library(ggplot2)
library(teal)
tm_plot <- function(label = "module") {
  module(
    label = label,
    ui = function(id) {
      ns <- NS(id)
      div(
        selectInput(ns("dataname"), label = "select dataname", choices = NULL),
        selectInput(ns("x"), label = "select x", choices = NULL),
        selectInput(ns("y"), label = "select y", choices = NULL),
        plotOutput(ns("plot")),
        verbatimTextOutput(ns("text"))
      )
    },
    server = function(id, data) {
      moduleServer(id, function(input, output, session) {
        observeEvent(data(), {
          updateSelectInput(inputId = "dataname", choices = names(data()))
        })
        observeEvent(input$dataname, {
          req(input$dataname)
          updateSelectInput(inputId = "x", choices = colnames(data()[[input$dataname]]))
          updateSelectInput(inputId = "y", choices = colnames(data()[[input$dataname]]))
        })
        dataname <- reactive(req(input$dataname))
        x <- reactive({
          req(input$x, input$x %in% colnames(data()[[dataname()]]))
          input$x
        })
        y <- reactive({
          req(input$y, input$y %in% colnames(data()[[dataname()]]))
          input$y
        })
        plot_data <- reactive({
          req(dataname(), x(), y())
          within(data(),
            {
              plot <- ggplot2::ggplot(dataname, ggplot2::aes(x = x, y = y)) +
                ggplot2::geom_point()
            },
            dataname = as.name(dataname()),
            x = as.name(x()),
            y = as.name(y())
          )
        })
        plot_r <- reactive({
          plot_data()[["plot"]]
        })
        output$plot <- renderPlot(plot_r())
        output$text <- renderText({
          teal.code::get_code(req(plot_data()))
        })
      })
    }
  )
}
app <- init(
  data = teal_data(iris = iris, mtcars = mtcars),
  modules = modules(
    tm_plot("identity")
  )
)
if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

```{r shinylive_iframe_1, echo = FALSE, out.width = '150%', out.extra = 'style = "position: relative; z-index:1"', eval = requireNamespace("roxy.shinylive", quietly = TRUE) && knitr::is_html_output() && identical(Sys.getenv("IN_PKGDOWN"), "true")}
code <- paste0(c(
  "interactive <- function() TRUE",
  knitr::knit_code$get("undecorated_app")
), collapse = "\n")
url <- roxy.shinylive::create_shinylive_url(code)
knitr::include_url(url, height = "800px")
```

#### Updating the module so it can be decorated

This module has one output, a plot created with `ggplot2`, and it displays the reproducible code used to obtain the plot.
We need to pass the `decorators` argument to the `ui` and `server` functions so they can be consumed to decorate the plot.
Note that we call the `ui_transform_teal_data` and `srv_transform_teal_data` functions to apply these decorators.

```{r tm_decorated_plot}
tm_decorated_plot <- function(label = "module", decorators = NULL) {
  checkmate::assert_list(decorators, "teal_transform_module", null.ok = TRUE)

  module(
    label = label,
    ui_args = list(decorators = decorators),
    server_args = list(decorators = decorators),
    ui = function(id, decorators) {
      ns <- NS(id)
      div(
        selectInput(ns("dataname"), label = "select dataname", choices = NULL),
        selectInput(ns("x"), label = "select x", choices = NULL),
        selectInput(ns("y"), label = "select y", choices = NULL),
        ui_transform_teal_data(ns("decorate"), transformators = decorators),
        plotOutput(ns("plot")),
        verbatimTextOutput(ns("text"))
      )
    },
    server = function(id, data, decorators) {
      moduleServer(id, function(input, output, session) {
        observeEvent(data(), {
          updateSelectInput(inputId = "dataname", choices = names(data()))
        })

        observeEvent(input$dataname, {
          req(input$dataname)
          updateSelectInput(inputId = "x", choices = colnames(data()[[input$dataname]]))
          updateSelectInput(inputId = "y", choices = colnames(data()[[input$dataname]]))
        })

        dataname <- reactive(req(input$dataname))
        x <- reactive({
          req(input$x, input$x %in% colnames(data()[[dataname()]]))
          input$x
        })
        y <- reactive({
          req(input$y, input$y %in% colnames(data()[[dataname()]]))
          input$y
        })

        # Plot is created within the teal_data object
        data_with_plot <- reactive({
          req(dataname(), x(), y())
          within(data(),
            {
              plot <- ggplot2::ggplot(dataname, ggplot2::aes(x = x, y = y)) +
                ggplot2::geom_point()
            },
            dataname = as.name(dataname()),
            x = as.name(x()),
            y = as.name(y())
          )
        })

        # Decorators are applied
        data_with_plot_decorated <- srv_transform_teal_data(
          "decorate",
          data = data_with_plot,
          transformators = decorators
        )

        # (Decorated) plot object is extracted for rendering
        plot_r <- reactive({
          data_with_plot_decorated()[["plot"]]
        })

        # Add plot printing statement to reproducible code
        ## This does not affect the analysis but when the code is "replayed"
        ## in an interactive session it will send the plot to a graphics device.
        reproducible_code <- reactive({
          within(data_with_plot_decorated(), expr = plot) |>
            teal.code::get_code()
        })

        output$plot <- renderPlot(plot_r())
        output$text <- renderText(reproducible_code())
      })
    }
  )
}
```


## Building Decorators

### Server

Here we create a simple decorator that does not provide user input.
Knowing that the module contains an object of class `ggplot2` named `plot`, we will modify its title and x-axis title:

```{r static_decorator}
static_decorator <- teal_transform_module(
  label = "Static decorator",
  server = function(id, data) {
    moduleServer(id, function(input, output, session) {
      reactive({
        req(data())
        within(data(), {
          plot <- plot +
            ggtitle("This is a better title") +
            xlab("the real x axis")
        })
      })
    })
  }
)
```


### UI

If decoration requires user input, a `ui` function can be added.
Here, the x-axis title is obtained from a `textInput` widget, giving the user some flexibility.
Note how the input values are passed to the `within()` function using its `...` argument.
See `?teal.code::within.qenv` for more examples.

```{r interactive_decorator}
interactive_decorator <- teal_transform_module(
  label = "Interactive decorator",
  ui = function(id) {
    ns <- NS(id)
    div(
      textInput(ns("x_axis_title"), "X axis title", value = "the suggested x axis")
    )
  },
  server = function(id, data) {
    moduleServer(id, function(input, output, session) {
      reactive({
        req(data())
        within(data(),
          {
            plot <- plot +
              ggtitle("This is a better title") +
              xlab(my_title)
          },
          my_title = input$x_axis_title
        )
      })
    })
  }
)
```


### Variable Names as Arguments

The server function of a decorating `teal_transform_module` must conform to the names of the variables that exist in the server function of the decorated `teal_module`.
Writing a universal decorator that applies to any module is impossible because different modules may use different variable names for their output elements.
It is possible, however, to create a decorator that will take the relevant variable names as arguments.
Here, the `output_name` variable name is passed to a decorator, allowing it to work with multiple modules.

```{r dynamic_decorator}
dynamic_decorator <- function(output_name) {
  teal_transform_module(
    label = "Dynamic decorator",
    ui = function(id) {
      ns <- NS(id)
      div(
        textInput(ns("x_axis_title"), "X axis title", value = "the syggested x axis")
      )
    },
    server = function(id, data) {
      moduleServer(id, function(input, output, session) {
        reactive({
          req(data())
          within(data(),
            {
              output_name <- output_name +
                xlab(x_axis_title)
            },
            output_name = as.name(output_name),
            x_axis_title = input$x_axis_title
          )
        })
      })
    }
  )
}
```

Note that when the function is used, `output_name` will be passed a character string but the expression passed to `within` needs a `name`/`symbol`, a language object, hence the argument value must be converted to a `name`.


## Using Decorators

Decorators are applied to a `teal` module as follows:

1. A list of decorators is passed to the module constructor function (_e.g._ `tm_my_module`).
2. The module constructor calls the module generator function (`teal::module`) and passes the decorators to the `ui_args` and `server_args` arguments.
3. The module functions, UI and server, take a list of decorators as arguments and resolve them using `ui_transform_teal_data` and `srv_transform_teal_data`, respectively.

The following examples demonstrate various uses of decorators.

### Single Decoration

In the first example we will apply one decoration to one output.

#### Application

Note that every call to the module constructor (`tm_decorated_plot`) takes a list containing _one_ decorator.

```{r app_2}
app <- init(
  data = teal_data(iris = iris, mtcars = mtcars),
  modules = modules(
    tm_decorated_plot("undecorated"),
    tm_decorated_plot("static", decorators = list(static_decorator)),
    tm_decorated_plot("interactive", decorators = list(interactive_decorator)),
    tm_decorated_plot("dynamic", decorators = list(dynamic_decorator("plot")))
  )
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

```{r shinylive_iframe_2, echo = FALSE, out.width = '150%', out.extra = 'style = "position: relative; z-index:1"', eval = requireNamespace("roxy.shinylive", quietly = TRUE) && knitr::is_html_output() && identical(Sys.getenv("IN_PKGDOWN"), "true")}
code <- paste0(c(
  "interactive <- function() TRUE",
  knitr::knit_code$get("setup"),
  knitr::knit_code$get("static_decorator"),
  knitr::knit_code$get("interactive_decorator"),
  knitr::knit_code$get("dynamic_decorator"),
  knitr::knit_code$get("tm_decorated_plot"),
  knitr::knit_code$get("app_2")
), collapse = "\n")
url <- roxy.shinylive::create_shinylive_url(code)
knitr::include_url(url, height = "800px")
```


### Decorating Multiple Outputs

Here we will apply decoration to two outputs in one module.

#### Decorators

The plot decorator adds a user-provided title to a `ggplot2` object.

```{r plot_decorator}
plot_decorator <- teal_transform_module(
  label = "Decorate plot",
  ui = function(id) {
    ns <- NS(id)
    textInput(ns("plot_title"), "Plot Title", value = "Title (editable)")
  },
  server = function(id, data) {
    moduleServer(id, function(input, output, session) {
      reactive({
        req(data())
        within(data(),
          {
            plot <- plot + ggplot2::ggtitle(ptitle) +
              ggplot2::theme_minimal() +
              ggplot2::theme(
                plot.title = element_text(face = "bold", size = 30, color = "blue")
              )
          },
          ptitle = input$plot_title
        )
      })
    })
  }
)
```

The table decorator adds a column to a `data.frame`.

```{r table_decorator}
table_decorator <- teal_transform_module(
  label = "Decorate table",
  ui = function(id) shiny::tags$p("No UI needed for table decorator and could be ommited."),
  server = function(id, data) {
    moduleServer(id, function(input, output, session) {
      reactive({
        req(data())
        within(data(), {
          table_data[["Added by decorator"]] <- paste0("Row ", seq_len(nrow(table_data)))
        })
      })
    })
  }
)
```


#### Module

The following module uses `ggplot2` to generate a scatter plot, and presents a simple `data.frame` as a summary table.
Code for both outputs is also displayed.

Note that the module constructor accepts one list of decorators and the decorators are then manually separated in the module functions.

```{r tm_decorated_plot_table}
tm_decorated_plot_table <- function(label = "module with two outputs", decorators = list()) {
  checkmate::assert_list(decorators, "teal_transform_module", null.ok = TRUE)

  module(
    label = label,
    ui_args = list(decorators = decorators),
    server_args = list(decorators = decorators),
    ui = function(id, decorators) {
      ns <- NS(id)
      div(
        selectInput(ns("dataname"), label = "Select dataset", choices = NULL),
        selectInput(ns("x"), label = "Select x-axis", choices = NULL),
        selectInput(ns("y"), label = "Select y-axis", choices = NULL),

        # Separately inject UI for plot and table decorators
        ui_transform_teal_data(ns("decorate_plot"), transformators = decorators$plot),
        ui_transform_teal_data(ns("decorate_table"), transformators = decorators$table),
        plotOutput(ns("plot")),
        tableOutput(ns("table")),
        verbatimTextOutput(ns("text"))
      )
    },
    server = function(id, data, decorators) {
      moduleServer(id, function(input, output, session) {
        observeEvent(data(), {
          updateSelectInput(inputId = "dataname", choices = names(data()))
        })

        dataname <- reactive(req(input$dataname))

        observeEvent(dataname(), {
          updateSelectInput(inputId = "x", choices = colnames(data()[[input$dataname]]))
          updateSelectInput(inputId = "y", choices = colnames(data()[[input$dataname]]))
        })
        x <- reactive({
          req(input$x, input$x %in% colnames(data()[[dataname()]]))
          input$x
        })
        y <- reactive({
          req(input$y, input$y %in% colnames(data()[[dataname()]]))
          input$y
        })

        # Separately create outputs within teal_data objects in separate reactive expressions
        plot_data <- reactive({
          req(dataname(), x(), y())
          within(data(),
            {
              plot <- ggplot2::ggplot(dataname, ggplot2::aes(x = xvar, y = yvar)) +
                ggplot2::geom_point()
            },
            dataname = as.name(dataname()),
            xvar = as.name(x()),
            yvar = as.name(y())
          )
        })
        table_data <- reactive({
          req(dataname())
          within(data(),
            {
              table_data <- data.frame(lapply(dataname, mean, na.rm = TRUE))
            },
            dataname = as.name(dataname())
          )
        })

        # Separately apply decoration to the outputs
        decorated_plot <- srv_transform_teal_data(
          "decorate_plot",
          data = plot_data,
          transformators = decorators$plot
        )
        decorated_table <- srv_transform_teal_data(
          "decorate_table",
          data = table_data,
          transformators = decorators$table
        )

        output$plot <- renderPlot(decorated_plot()[["plot"]])
        output$table <- renderTable(decorated_table()[["table_data"]])

        output$text <- renderText({
          plot_code <- teal.code::get_code(req(decorated_plot()))
          table_code <- teal.code::get_code(req(decorated_table()))
          paste("# Plot Code:", plot_code, "\n\n# Table Code:", table_code)
        })
      })
    }
  )
}
```


#### Application

Note that a named list of decorators is passed to the module constructor.

```{r app_3}
app <- init(
  data = teal_data(iris = iris, mtcars = mtcars),
  modules = modules(
    tm_decorated_plot_table(
      "plot_and_table",
      decorators = list(
        plot = plot_decorator,
        table = table_decorator
      )
    )
  )
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

```{r shinylive_iframe_3, echo = FALSE, out.width = '150%', out.extra = 'style = "position: relative; z-index:1"', eval = requireNamespace("roxy.shinylive", quietly = TRUE) && knitr::is_html_output() && identical(Sys.getenv("IN_PKGDOWN"), "true")}
code <- paste0(c(
  "interactive <- function() TRUE",
  knitr::knit_code$get("setup"),
  knitr::knit_code$get("plot_decorator"),
  knitr::knit_code$get("table_decorator"),
  knitr::knit_code$get("tm_decorated_plot_table"),
  knitr::knit_code$get("app_3")
), collapse = "\n")
url <- roxy.shinylive::create_shinylive_url(code)
knitr::include_url(url, height = "800px")
```


## Convenience

Here we present some ways to work with decorators more conveniently.
These are purely optional.

### Reducing Boilerplate

The function `make_teal_transform_server` can be used to reduce the amount of boilerplate code when writing new decorators.
It takes `language` as input and requires you to use `input` object names directly in the expression.
The following calls yield the same decorator module.
Note that the combination of `my_title = input$x_axis_title` and `xlab(my_title)` is replaced by a simple `xlab(x_axis_table)`.
```{r, eval=FALSE}
teal_transform_module(
  label = "Static decorator",
  ui = function(id) {
    ns <- NS(id)
    div(
      textInput(ns("x_axis_title"), "X axis title", value = "x axis")
    )
  },
  server = function(id, data) {
    moduleServer(id, function(input, output, session) {
      reactive({
        req(data())
        within(
          data(),
          {
            plot <- plot + ggtitle("This is a better title") + xlab(x_axis_title)
          },
          x_axis_title = input$x_axis_title
        )
      })
    })
  }
)

teal_transform_module(
  label = "Static decorator (language)",
  ui = function(id) {
    ns <- NS(id)
    div(
      textInput(ns("x_axis_title"), "X axis title", value = "x axis")
    )
  },
  server = make_teal_transform_server(
    expression(
      plot <- plot + ggtitle("This is a better title") + xlab(x_axis_title)
    )
  )
)
```


### Multiple Decorators

Consider these constructs to accommodate an arbitrary number of decorators in your module.
Note that with this method all decorations will be applied to one output.
```{r, eval=FALSE}
# in the module UI function
div(
  id = ns("deorator_container"),
  lapply(names(decorators), function(decorator_name) {
    div(
      id = ns(paste0("decorate_", decorator_name)),
      ui_transform_teal_data(
        ns(paste0("decorate_", decorator_name)),
        transformators = decorators[[decorator_name]]
      )
    )
  })
)
# in the module server function
output_data <- reactive(teal_data())
decorations <- lapply(names(decorators), function(decorator_name) {
  function(data) {
    srv_transform_teal_data(
      paste0("decorate_", decorator_name),
      data = data,
      transformators = decorators[[decorator_name]]
    )
  }
})
output_data_decorated <- Reduce(function(f, ...) f(...), decorations, init = output_data, right = TRUE)
```
