---
title: "Data Transformations as Shiny Module"
author: "NEST CoreDev"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Data Transformations as shiny Module}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

`teal_transform_module()` is a Shiny module that takes `ui` and `server` arguments. When provided, `teal` executes data transformations for the specified module both when it is loaded and whenever the data changes.
The `server` argument extends the logic behind data manipulations, while the `ui` argument adds new UI elements to the filter panel, managing the transformator inputs.

In version `0.16`, `teal` introduced a new optional argument in `teal::module` named `transformators`.
This argument allows users to pass a `list` of objects of `teal_data_module` class, created using the `teal_transform_module()` function.

The main advantage of `teal_transform_module()` is its ability to transform data before passing it to the module.
This feature allows users to extend the default behavior of existing modules by performing specific data operations on the data within the module.

This vignette presents the way on how to manage custom data transformations in `teal` apps.

## Creating Your First Custom Data Transformation Module

Let's initialize a simple `teal` app by providing `iris` and `mtcars` as input datasets.

```{r, message = FALSE, warning = FALSE}
library(teal)
```

```{r}
data <- within(teal_data(), {
  iris <- iris
  mtcars <- mtcars
})

app <- init(
  data = data,
  modules = teal::example_module()
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

### Single transformator

Next, let's create a **single transformator** using `teal_transform_module()`.
This example enables users to display the first `n` rows of the `iris` dataset based on their input.

Here’s how we can implement this:

1. **UI Component**: We create a `numericInput` in the `ui` that allows users to specify the number of rows they want to display.
2. **Server Logic**: In the `server` function, we take in a reactive `data` object, perform the transformation to extract the first `n` rows of the `iris` dataset, and return the updated reactive `data`.

```{r}
data <- within(teal_data(), {
  iris <- iris
  mtcars <- mtcars
})

my_transformators <- list(
  teal_transform_module(
    label = "Custom transformator for iris",
    ui = function(id) {
      ns <- NS(id)
      tags$div(
        numericInput(
          ns("n_rows"),
          "Number of rows to subset",
          value = 6, min = 1, max = 150, step = 1
        )
      )
    },
    server = function(id, data) {
      moduleServer(id, function(input, output, session) {
        reactive({
          within(data(),
            {
              iris <- head(iris, num_rows)
            },
            num_rows = input$n_rows
          )
        })
      })
    }
  )
)

app <- init(
  data = data,
  modules = teal::example_module(transformators = my_transformators)
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

> **_Note_**: It is recommended to return `reactive()` with `teal_data()` in the `server` code of a `teal_transform_module`, as this approach ensures better robustness in maintaining Shiny's reactivity.
If you are planning on using `eventReactive()` in the `server`, make sure the event includes `data()` _(example `eventReactive(list(input$a, data()), {...})`)_.
More in [this discussion](https://github.com/insightsengineering/teal/issues/1303#issuecomment-2286239832).

### Multiple transformators

We could also add **multiple transformators** to the app by including several instances of `teal_transform_module` in a list.

For example, let’s add another transformation to the `mtcars` dataset. This transformation creates a new column containing the `rownames` of `mtcars`.
Unlike the previous example, this module does not include any interactive UI elements.

```{r}
data <- within(teal_data(), {
  iris <- iris
  mtcars <- mtcars
})

my_transformators <- list(
  teal_transform_module(
    label = "Custom transformator for iris",
    ui = function(id) {
      ns <- NS(id)
      tags$div(
        numericInput(
          ns("n_rows"),
          "Number of rows to subset",
          value = 6, min = 1, max = 150, step = 1
        )
      )
    },
    server = function(id, data) {
      moduleServer(id, function(input, output, session) {
        reactive({
          within(data(),
            {
              iris <- head(iris, num_rows)
            },
            num_rows = input$n_rows
          )
        })
      })
    }
  ),
  teal_transform_module(
    label = "Custom transformator for mtcars",
    ui = function(id) {
      ns <- NS(id)
      tags$div(
        "Adding rownames column to mtcars"
      )
    },
    server = function(id, data) {
      moduleServer(id, function(input, output, session) {
        reactive({
          within(data(), {
            mtcars$rownames <- rownames(mtcars)
            rownames(mtcars) <- NULL
          })
        })
      })
    }
  )
)

app <- init(
  data = data,
  modules = teal::example_module(transformators = my_transformators)
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```
