---
title: "Data as shiny Module"
author: "NEST CoreDev"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Data as shiny Module}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

Proper functioning of any `teal` application requires presence of a `teal_data` object.
Typically, a `teal_data` object created in the global environment will be passed to the `data` argument in `init`.
This `teal_data` object should contain all elements necessary for successful execution of the application's modules. 

In some scenarios, however, application developers may opt to postpone some data operations until the application run time.
This can be done by passing a special _`shiny` module_ to the `data` argument.
The `teal_data_module` function is used to build such a module from the following components:

- a `ui` function; accepts only one argument, `id`; defines user interface elements for the data module
- a `server` function: accepts only one argument, `id`; defines server logic for the data module, including data creation; must return a reactive expression containing a `teal_data` object

`teal` will run this module when the application starts and the resulting `teal_data` object that will be used throughout all `teal` (analytic) modules. 

## Creating Data In-App

One case for postponing data operations are data sets that are dynamic, frequently updated.
Such data cannot be created once and kept in the global environment.
Using `teal_data_module` allows to create a data set from scratch every time the user starts the application.

```{r, message = FALSE, warning = FALSE}
library(teal)
```


```{r}
data_mod <- teal_data_module(
  ui = function(id) div(),
  server = function(id) {
    moduleServer(id, function(input, output, session) {
      reactive({
        data <- within(
          teal_data(),
          {
            dataset1 <- iris
            dataset2 <- mtcars
          }
        )
        datanames(data) <- c("dataset1", "dataset2")
        data
      })
    })
  }
)


app <- init(
  data = data_mod,
  module = example_module()
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```


## Modification of Data In-App

Another reason to postpone data operations is to allow the application user to act the preprocessing stage.
An initial, constant form of the data can be created in the global environment and then modified once the app starts.

The following example illustrates how `teal_data_module` can be utilized to subset data based on the user inputs:

```{r}
data <- within(teal_data(), {
  dataset1 <- iris
  dataset2 <- mtcars
})
datanames(data) <- c("dataset1", "dataset2")

data_mod <- teal_data_module(
  ui = function(id) {
    ns <- NS(id)
    div(
      selectInput(ns("species"), "Select species to filter",
        choices = unique(iris$Species), multiple = TRUE
      ),
      actionButton(ns("submit"), "Submit")
    )
  },
  server = function(id) {
    moduleServer(id, function(input, output, session) {
      eventReactive(input$submit, {
        data_modified <- within(
          data,
          dataset1 <- subset(dataset1, Species %in% selected),
          selected = input$species
        )
        data_modified
      })
    })
  }
)

app <- init(
  data = data_mod,
  module = example_module()
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

_See `?within.qenv` for a detailed explanation of how to use the `within` method._

Note that running preprocessing code in a module as opposed to the global environment will increase app loading times.
It is recommended to keep the constant code in the global environment and to move only the dynamic parts to a data module.

###### WARNING

When using `teal_data_module` to modify a pre-existing `teal_data` object it is crucial that the server function and the data object are defined in the same environment as otherwise the server function will not be able to access the data object.
This means server functions defined in packages cannot be used.

### Extending existing `teal_data_modules`

The `teal_data_module` can be further modified outside of the initial `shiny` module and processed after user inputs.
The `within` function allows to process at runtime the data in the `teal_data` object contained in `teal_data_module`.

Building on the previous example, the `data_mod` is handled as a generic connector and here new columns are added once the data is retrieved.

```{r}
data_mod_2 <- within(
  data_mod,
  {
    # Create new column with Ratio of Sepal.Width and Petal.Width
    dataset1$Ratio.Sepal.Petal.Width <- round(dataset1$Sepal.Width / dataset1$Petal.Width, digits = 2L)
    # Create new column that converts Miles per Galon to Liter per 100 Km
    dataset2$lp100km <- round(dataset2$mpg * 0.42514371, digits = 2L)
    dataset2 <- dplyr::relocate(
      dataset2,
      "lp100km",
      .after = "mpg"
    )
  }
)

app <- init(
  data = data_mod_2,
  module = example_module()
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```
