---
title: "Decorate modules output"
author: "NEST CoreDev"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Decorate modules output}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Create a decoration object

`teal_transform_module` without `ui`. Nothing will be displayed, only server code will be executed.

```{r}
library(teal)
static_decorator <- teal_transform_module(
  server = function(id, data) {
    moduleServer(id, function(input, output, session) {
      reactive({
        req(data())
        data() |> within({
          plot <- plot +
            ggtitle("This is title") +
            xlab("x axis")
        })
      })
    })
  }
)
```

Use `make_teal_transform_server(expr)` to create a transformer without repeating a boiler plate code (`function(id, data)`, `moduleServer`, `reactive`, `within(data, ...)`)

```{r}
static_decorator_lang <- teal_transform_module(
  server = make_teal_transform_server(
    expression(
      plot <- plot +
        ggtitle("This is title") +
        xlab("x axis title")
    )
  )
)
```

Adding interactivity requires creating a proper shiny module.
```{r}
interactive_decorator <- teal_transform_module(
  ui = function(id) {
    ns <- NS(id)
    div(
      textInput(ns("x_axis_title"), "X axis title", value = "x axis")
    )
  },
  server = function(id, data) {
    moduleServer(id, function(input, output, session) {
      reactive({
        req(data())
        data() |> within(
          {
            plot <- plot +
              ggtitle("This is title") +
              xlab(x_axis_title)
          },
          x_axis_title = input$x_axis_title
        )
      })
    })
  }
)
```

There is `make_teal_transform_server(expr)` which simplifies a creation of `teal_transform_module`'s server. One can 
simply specify `expression` within which object names can be replaced with respective input.

```{r}
interactive_decorator_lang <- teal_transform_module(
  ui = function(id) {
    ns <- NS(id)
    div(
      textInput(ns("x_axis_title"), "X axis title", value = "x axis")
    )
  },
  server = make_teal_transform_server(
    expression(
      plot <- plot +
        ggtitle("This is title") +
        xlab(x_axis_title)
    )
  )
)
```


Since `teal_transform_module` when created has constant object names it means that it is hard to robustly reuse
transformer in any `tm_` module. It is recomended for external parties to collect library of "decorators" as functions
to potentially adjust the content of the evaluated expression to `teal_modules`'s internals. See the following example
and focus on `output_name`.

    ```{r}
gg_xlab_decorator <- function(output_name) {
  teal_transform_module(
    ui = function(id) {
      ns <- NS(id)
      div(
        textInput(ns("x_axis_title"), "X axis title", value = "x axis")
      )
    },
    server = function(id, data) {
      moduleServer(id, function(input, output, session) {
        reactive({
          req(data())
          data() |> within(
            {
              output_name <- output_name +
                xlab(x_axis_title)
            },
            x_axis_title = input$x_axis_title,
            output_name = as.name(output_name)
          )
        })
      })
    }
  )
}
    ```


Failures in decorators are handled by internal `teal` mechanism called `trigger_on_success`, which will never 
break `data` object inside of the module. Instead decorator will be ignored and relevant error message will be shown.

```{r}
failing_decorator <- teal_transform_module(
  ui = function(id) {
    ns <- NS(id)
    div(
      textInput(ns("x_axis_title"), "X axis title", value = "x axis")
    )
  },
  server = function(id, data) {
    moduleServer(id, function(input, output, session) {
      reactive(stop("Hihi"))
    })
  }
)
    ```

## Example module

To include decorators to the `teal_module` one needs to add an extra argument to the module-constructor function `tm_`. Decorators should be passed via `ui_args` and `server_args` to the module's `ui` and `server`, where they should be 
consumed by `ui/srv_teal_transform_module`.


```{r}
tm_decorated_plot <- function(label = "module", transforms = list(), decorators = teal_transform_module()) {
  module(
    label = label,
    ui = function(id, decorators) {
      ns <- NS(id)
      div(
        selectInput(ns("dataname"), label = "select dataname", choices = NULL),
        selectInput(ns("x"), label = "select x", choices = NULL),
        selectInput(ns("y"), label = "select y", choices = NULL),
        ui_teal_transform_module(ns("decorate"), transforms = decorators),
        plotOutput(ns("plot")),
        verbatimTextOutput(ns("text"))
      )
    },
    server = function(id, data, decorators) {
      moduleServer(id, function(input, output, session) {
        observeEvent(data(), {
          updateSelectInput(inputId = "dataname", choices = teal.data::datanames(data()))
        })

        observeEvent(input$dataname, {
          updateSelectInput(inputId = "x", choices = colnames(data()[[input$dataname]]))
          updateSelectInput(inputId = "y", label = "select y", choices = colnames(data()[[input$dataname]]))
        })

        q1 <- reactive({
          req(input$dataname, input$x, input$y)
          data() |>
            within(
              {
                plot <- ggplot2::ggplot(dataname, ggplot2::aes(x = x, y = y)) +
                  ggplot2::geom_point()
              },
              dataname = as.name(input$dataname),
              x = as.name(input$x),
              y = as.name(input$y)
            )
        })

        q2 <- srv_teal_transform_module("decorate", data = q1, transforms = decorators)

        plot_r <- reactive({
          req(q2())
          q2()[["plot"]]
        })

        output$plot <- renderPlot(plot_r())
        output$text <- renderText({
          teal.code::get_code(q2())
        })
      })
    },
    ui_args = list(decorators = decorators),
    server_args = list(decorators = decorators)
  )
}
```

## Example app

```{r}
library(ggplot2)
app <- init(
  data = teal_data(iris = iris, mtcars = mtcars),
  modules = modules(
    tm_decorated_plot("identity"),
    tm_decorated_plot("no-ui", decorator = static_decorator),
    tm_decorated_plot("lang", decorator = static_decorator_lang),
    tm_decorated_plot("interactive", decorator = interactive_decorator),
    tm_decorated_plot("interactive-from lang", decorator = interactive_decorator_lang),
    tm_decorated_plot("from-fun", decorator = gg_xlab_decorator("plot")),
    tm_decorated_plot("failing", decorator = failing_decorator)
  )
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

## To improve

- Error handling when a decorator throws an error.
- Rewrite modules so that they always have a graph as `graph` (or `plot`, `g`, `p` etc.) and table as 
  `table` (or `t` etc.) and listings as `listings` (or `l` etc.). This simplifies/standardizes a construction of
  `teal_transform_module` which can be re-used in multiple modules. 