---
title: "Using ADaM Data in teal Applications"
author: "NEST CoreDev"
date: "2022-04-20"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using ADaM Data in teal Applications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

To include `ADaM` data in a `teal` app, the `teal.data::cdisc_data` function is used.

`cdisc_data` allows `teal` applications to include multiple datasets, identifying merge keys and providing information to produce R code for reproducibility.

There is an advantage to passing `CDISC` datasets that adhere to `ADaM` standards to these functions in that the code is minimized. However, the dataset-related functions also include the flexibility to work with non-standard datasets provided that merge keys and the relationship between the datasets are specified.

The examples below illustrate the usage of these different dataset functions for example `cdisc_dataset` and `dataset`. For more information, see documentation in `teal.data`.

## Keys

Primary keys serve as unique row identifiers in individual datasets and thus need to be specified for each dataset and dataset connector. These can be specified on the most general dataset constructor, `dataset`, as shown below.

```{r, message=FALSE}
library(teal)

# using cdisc_dataset, keys are automatically derived for standard datanames
# (although they can be overwritten)
adsl <- data.frame(
  STUDYID = "study",
  USUBJID = 1:10,
  SEX = sample(c("F", "M"), 10, replace = TRUE),
  AGE = rpois(10, 40)
)
dataset_adsl <- cdisc_dataset("ADSL", adsl)
class(dataset_adsl)
```

When passing multiple datasets to the `cdisc_data` function, dataset relationship are set using `join_keys` and `join_key` and these are used to merge datasets together within `teal` apps.

In the example below, two standard `CDISC` datasets (`ADSL` and `ADTTE`) are passed to `cdisc_data`. In the case of `CDISC` datasets that adhere to `ADaM` standards, the merge keys do not need to be manually specified. Keys are automatically added if `dataname` matches one of the implemented standards as documented in the `cdisc_dataset` function. This minimizes the code needed to allow data merges as seen in this example:

```{r, message=FALSE}
adsl <- data.frame(
  STUDYID = "study",
  USUBJID = 1:10,
  SEX = sample(c("F", "M"), 10, replace = TRUE),
  AGE = rpois(10, 40)
)
adtte <- rbind(adsl, adsl, adsl)
adtte$PARAMCD <- rep(c("OS", "EFS", "PFS"), each = 10)
adtte$AVAL <- c(
  rnorm(10, mean = 700, sd = 200), # dummy OS level
  rnorm(10, mean = 400, sd = 100), # dummy EFS level
  rnorm(10, mean = 450, sd = 200) # dummy PFS level
)

cdisc_data_obj <- cdisc_data(
  cdisc_dataset(
    dataname = "ADSL",
    x = adsl,
    code = '
    adsl <- data.frame(
      STUDYID = "study",
      USUBJID = 1:10,
      SEX = sample(c("F", "M"), 10, replace = TRUE),
      AGE = rpois(10, 40)
    )'
  ),
  cdisc_dataset(
    dataname = "ADTTE",
    x = adtte,
    code = '
    adtte <- rbind(adsl, adsl, adsl)
    adtte$PARAMCD <- rep(c("OS", "EFS", "PFS"), each = 10)
    adtte$AVAL <- c(
      rnorm(10, mean = 700, sd = 200),
      rnorm(10, mean = 400, sd = 100),
      rnorm(10, mean = 450, sd = 200)
    )'
  )
)
class(cdisc_data_obj)
```

which is equivalent to:
```{r, message=FALSE}
example_data <- cdisc_data(
  cdisc_dataset(
    dataname = "ADSL",
    x = adsl,
    code = '
    adsl <- data.frame(
      STUDYID = "study",
      USUBJID = 1:10,
      SEX = sample(c("F", "M"), 10, replace = TRUE),
      AGE = rpois(10, 40)
    )',
    keys = c("STUDYID", "USUBJID")
  ),
  cdisc_dataset(
    dataname = "ADTTE",
    x = adtte,
    code = '
    adtte <- rbind(adsl, adsl, adsl)
    adtte$PARAMCD <- rep(c("OS", "EFS", "PFS"), each = 10)
    adtte$AVAL <- c(
      rnorm(10, mean = 700, sd = 200),
      rnorm(10, mean = 400, sd = 100),
      rnorm(10, mean = 450, sd = 200)
    )',
    keys = c("STUDYID", "USUBJID", "PARAMCD")
  ),
  join_keys = join_keys(
    join_key("ADSL", "ADSL", c("STUDYID", "USUBJID")),
    join_key("ADTTE", "ADTTE", c("USUBJID", "STUDYID", "PARAMCD")),
    join_key("ADSL", "ADTTE", c("STUDYID", "USUBJID"))
  )
)
class(cdisc_data_obj)
```

The `teal.data::join_keys` function is used to specify keys:

- [`teal.data::join_keys`](https://insightsengineering.github.io/teal.data/latest-tag/reference/join_keys.html) is a collection of multiple `teal.data::join_key` entries
- [`teal.data::join_key`](https://insightsengineering.github.io/teal.data/latest-tag/reference/join_key.html) specifies the relation between two datasets:
  - `dataset_1`, `dataset_2` - name of two datasets
  - `key` - optionally named vector of column names

Note that it is assumed that join keys are symmetric, i.e. `join_key("x", "y", "x_col" = "y_col")` will enable merge from "x" to "y" and vice-versa.


## Further Reading

For more information about preprocessing, reproducibility, relationships between datasets and `DDL`, please refer to the [`teal.data` package](https://insightsengineering.github.io/teal.data/).
