---
title: "Input general data in a teal application"
author: "NEST CoreDev"
date: "2022-04-20"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Input general data in a teal application}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

Teal applications are not restricted to `CDISC`-standard data. Although many teal modules included with `NEST` are designed for `CDISC` data, those in the library `teal.modules.general` have been designed to work with non-relational data.

For example this application uses the standard `iris` and `mtcars` datasets:

```{r, message=FALSE}
library(teal)

app <- init(
  data = teal_data(
    dataset("IRIS", iris, code = "IRIS <- iris"),
    dataset("CARS", mtcars, code = "CARS <- mtcars")
  ),
  modules = example_module()
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

For more information, see documentation in `teal.data`.

## Delayed Data Loading (`DDL`)

To learn more about `DDL`, visit `vignette("delayed-data-loading", package = "teal.data")`. All the features of `DDL`
are available for general datasets:

```{r eval=FALSE}
library(teal)

person_generator <- function() {
  return(
    data.frame(
      ID = factor(1:8),
      AGE = c(40, 23, 56, 11, 17, 71, 23, 56)
    )
  )
}

pet_generator <- function() {
  return(
    data.frame(
      ID = factor(1:10),
      TYPE = rep(c("CAT", "DOG"), 5),
      COLOR = c("GINGER", rep("BROWN", 5), rep("BLACK", 4)),
      PERSON_ID = factor(c(5, 4, 3, 3, 3, 1, 8, 1, 2, 2))
    )
  )
}


app <- init(
  data = teal_data(
    fun_dataset_connector("PERSON", fun = person_generator, keys = "ID") %>%
      mutate_dataset("PERSON$SEX <- rep(c('M', 'F'), 4)"),
    fun_dataset_connector("PETS", fun = pet_generator, keys = "ID")
  ) %>%
    mutate_join_keys("PERSON", "PETS", c("ID" = "PERSON_ID")),
  modules = example_module()
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

For more information about preprocessing, reproducibility, relationships between datasets and `DDL`, please refer to the [`teal.data` package](https://insightsengineering.github.io/teal.data/).
