---
title: "Including Data in teal Applications"
author: "NEST CoreDev"
output: 
  bookdown::html_document2:
    base_format: rmarkdown::html_vignette
    toc: true
    toc_depth: 5
    number_sections: false
vignette: >
  %\VignetteIndexEntry{Including Data in teal Applications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Data in `teal` Applications

While the `teal` framework is mainly geared towards clinical data that conforms to the `ADaM` standard, general, non-relational data is handled just as well and the mechanism of passing data to applications is virtually the same.
Also, modules in `teal.modules.general` have been designed to work with general data.

All applications use the `teal_data` class as a data container.
`teal_data` objects are passed to `init` to build the application, where they are modified by the filter panel (if applicable) and passed on to modules.
Thus, the first step of building a `teal` app is creating a `teal_data` object.

### General Data

A `teal_data` object is created by calling the `teal_data` function and passing data objects as `name:value` pairs.

```{r, message=FALSE}
library(teal)

# create teal_data
data <- teal_data(iris = iris, cars = mtcars)
```

Note that `iris` and `cars` have been added to the `datanames` property of `data` (jump to [`teal_data` properties]).

This is sufficient to run a `teal` app.

```{r, eval=FALSE}
# build app
app <- init(
  data = data,
  modules = example_module()
)

# run app
shinyApp(app$ui, app$server)
```

### Reproducible Data

A `teal_data` stores data in a separate environment.
Therefore, modifying the stored datasets requires that code be evaluated in that environment.
Following that logic, one can create an empty `teal_data` object and populate it by evaluating code.
This can be done using the `eval_code` function or, more conveniently, using the `within` function.

```{r}
# create empty object
data0 <- teal_data()

# run code in the object
data1 <- teal.code::eval_code(data0, code = "iris <- iris
                                             cars <- mtcars"
)
# alternative
data2 <- within(data0, {
  iris <- iris
  cars <- mtcars
})
```

The key difference between `eval_code` and `within` is that the former accepts code as character vector or language objects (calls and expressions), while `within` accepts _only_ inline code.
See `?eval_code` for more details.

Note that with `data` the code that created the data objects is unknown so the process cannot be reproduced.
The necessary code can be added with the `code` argument of the `teal_data` function but in that case the code may not reproduce the environment.
Such an object is considered _unverified_ (jump to [`teal_data` properties]).

If reproducibility is required, we recommend creating `teal_data` empty and evaluating code.

###### code from file

The ability to pass code as character vector to `eval_code` opens the door to using code stored in a file.
```{r, eval=FALSE}
# not run
data_from_file <- teal_data()
data_from_file <- eval_code(data, readLines("<path_to_file>"))
```

### Creating Data In-App

The one departure from passing a `teal_data` object to `init` is when the data does not exist in the environment where the app is run, _e.g._ when it has to be pulled from a remote source.
In those cases a `teal_data_module` must be used.
See [this vignette](data-as-shiny-module.html) for a detailed description.

<br>

## Clinical Data

Currently `teal` supports two specialized data formats.

### `ADaM` Data

The `ADaM` data model specifies relationships between `CDISC` datasets.
The `cdisc_data` function takes advantage of that facto to automatically set default joining keys (jump to [`teal_data` properties]).
In the example below, two standard `CDISC` datasets (`ADSL` and `ADTTE`) are passed to `cdisc_data`.

```{r}
# create cdisc_data
data_cdisc <- cdisc_data(ADSL = teal.data::rADSL, ADTTE = teal.data::rADSL)

datanames(data_cdisc)
join_keys(data_cdisc)
```


```{r, eval=FALSE}
app <- init(
  data = data_cdisc,
  modules = example_module()
)
shinyApp(app$ui, app$server)
```

### `MultiAssayExperiment` Data

The `MultiAssayExperiment` package offers a data structure for representing and analyzing multi-omics experiments: a biological analysis approach utilizing multiple types of observations, such as DNA mutations and abundance of RNA and proteins, in the same biological specimens.

The `MultiAssayExperiment` class is described in detail [here](https://www.bioconductor.org/packages/release/bioc/vignettes/MultiAssayExperiment/inst/doc/MultiAssayExperiment.html).

`MultiAssayExperiment` objects (MAEs) are placed in `teal_data` just like normal objects.

```{r, eval=FALSE}
library(MultiAssayExperiment)
utils::data(miniACC)

data_mae <- teal_data(MAE = miniACC)

app <- init(
  data = data_mae,
  modules = example_module()
)
shinyApp(app$ui, app$server)
```

Note that due to the unique structure of a MAE `teal` requires special considerations when building `teal` modules.
Therefore, we cannot guarantee that all modules will work properly with MAEs.
The package [`teal.modules.hermes`](https://insightsengineering.github.io/teal.modules.hermes/latest-tag/) has been developed specifically with MAE in mind and will be more reliable.

The filter panel supports MAEs out of the box.

<br>

## `teal_data` properties

##### `datanames`

The `datanames` property lists the objects stored in the `teal_data` that constitute datasets of interest.
Objects passed to `teal_data` become automatically listed in the `datanames` property of the resulting object.
Objects created in `teal_data` by evaluating code need not be data objects of interest and as such they are not automatically added to `datanames`.
Use the `datanames` function to modify the `datanames` property.

```{r}
data_with_objects <- teal_data(iris = iris, cars = mtcars)
data_with_code <- teal_data() %>% 
  within({
    iris <- iris
    cars <- mtcars
    not_a_dataset <- "data source credits"
  })
datanames(data_with_objects)
datanames(data_with_code)
datanames(data_with_code) <- c("iris", "cars")
datanames(data_with_code)
```

All `teal` modules take a `datanames` argument that determines which datasets they are to have access to.
Only objects enumerated in the `datanames` property of the `teal_data` object can be used.

For a detailed explanation of datanames, see [this `teal.data` vignette](https://insightsengineering.github.io/teal.data/latest-tag/articles/teal-data.html).

(back to [General Data])

##### `join_keys`

Using relational data requires specifying joining keys for each pair of datasets.
Primary keys are unique row identifiers in individual datasets and thus should be specified for each dataset.
Foreign keys describe mapping of variables between datasets.
Joining keys are stored in the `join_keys` property, which can be set when creating a `teal_data` object, using the `join_keys` argument, or using the `join_keys` function.

```{r}
ds1 <- data.frame(
  id = seq(1, 10),
  group = rep(c("A", "B"), each = 5)
)
ds2 <- data.frame(
  group = c("A", "B"),
  condition = c("condition1", "condition2")
)
keys <- join_keys(
  join_key("DS1", keys = "id"),
  join_key("DS2", keys = "group"),
  join_key("DS1", "DS2", keys = c("group" = "group"))
)
data_relational1 <- teal_data(DS1 = ds1, DS2 = ds2, join_keys = keys)
data_relational2 <- teal_data(DS1 = ds1, DS2 = ds2)
join_keys(data_relational2) <- keys
```

For a detailed explanation of join keys, see [this `teal.data` vignette](https://insightsengineering.github.io/teal.data/latest-tag/articles/join-keys.html).

(back to [`ADaM` Data])

##### `verified`

`teal_data` allows for tracking code from data creation through filtering through analysis so that the whole process can be reproduced.
The `verified` property designates whether or not reproducibility has been confirmed.
`teal_data` objects that are created empty and are only modified by evaluating code in them are verified by default.
Ones created with data objects alone or with data objects and code are not verified by default but can become verified by running the `verify` function.

```{r}
data_with_code

data_with_objects_and_code <- teal_data(iris = iris, cars = mtcars, code = expression(iris <- iris, cars <- mtcars))
data_with_objects_and_code

data_with_objects_and_code_verified <- verify(data_with_objects_and_code)
data_with_objects_and_code_verified
```


For a detailed explanation of verification, see [this `teal.data` vignette](https://insightsengineering.github.io/teal.data/latest-tag/articles/teal-data-reproducibility.html).

(back to [Reproducible Data])

<br>

## Further Reading

For a complete guide to the `teal_data` class, please refer to the [`teal.data` package](https://insightsengineering.github.io/teal.data/latest-tag/).
