---
title: "Input ADaM data in a teal application"
author: "NEST CoreDev"
date: "2022-04-20"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Input ADaM data in a teal application}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

To include `ADaM` data in a teal app, the `teal.data::cdisc_data` function is used.

The `cdisc_data` function allows `teal` applications to include multiple datasets, identifying merge keys
and providing information to produce R code for reproducibility.

There is an advantage to passing `CDISC` datasets that adhere to `ADaM` standards to these functions in that the code is
minimized. However, the dataset-related functions also include the flexibility to work with non-standard datasets
provided that merge keys and the relationship between the datasets are specified.

The examples below illustrate the usage of these different dataset functions for example `cdisc_dataset` and `dataset`. For more information,
see documentation in `teal.data`.

## Keys

Primary keys serve as unique row identifiers in individual datasets and thus need to be specified for each dataset
and dataset connector. These can be specified on the most general dataset constructor `dataset` as
shown below.

```{r, message=FALSE}
library(teal)
library(scda)

# using cdisc_dataset, keys are automatically derived for standard datanames
# (although they can be overwritten)
adsl <- synthetic_cdisc_data("latest")$adsl
dataset_adsl <- cdisc_dataset("ADSL", adsl)
class(dataset_adsl)
```

When passing multiple datasets to the `cdisc_data` function, dataset relationship are set using
`join_keys` and `join_key` and these are used to merge datasets together within teal apps.

In the example below, two standard CDISC datasets (ADSL and ADTTE) are passed to the aforementioned
function. In the case of `CDISC` datasets that adhere to `ADaM` standards, the merge keys do not need to be manually
specified. Keys are automatically added if `dataname` matches one of the implemented standards as documented in the
`cdisc_dataset` function. This minimizes the code needed to allow data merges as seen in this
example:

```{r, message=FALSE}
adsl <- synthetic_cdisc_data("latest")$adsl
adtte <- synthetic_cdisc_data("latest")$adtte

cdisc_data_obj <- cdisc_data(
  cdisc_dataset(dataname = "ADSL", x = adsl, code = "ADSL <- synthetic_cdisc_data(\"latest\")$adsl"),
  cdisc_dataset(dataname = "ADTTE", x = adtte, code = "ADTTE <- synthetic_cdisc_data(\"latest\")$adtte")
)
class(cdisc_data_obj)

# which is equivalent to:
example_data <- cdisc_data(
  cdisc_dataset(
    dataname = "ADSL",
    x = adsl,
    code = "ADSL <- synthetic_cdisc_data(\"latest\")$adsl",
    keys = c("STUDYID", "USUBJID")
  ),
  cdisc_dataset(
    dataname = "ADTTE",
    x = adtte,
    code = "ADTTE <- synthetic_cdisc_data(\"latest\")$adtte",
    parent = "ADSL",
    keys = c("USUBJID", "STUDYID", "PARAMCD")
  ),
  join_keys = join_keys(
    join_key("ADSL", "ADSL", c("STUDYID", "USUBJID")),
    join_key("ADTTE", "ADTTE", c("USUBJID", "STUDYID", "PARAMCD")),
    join_key("ADSL", "ADTTE", c("STUDYID", "USUBJID"))
  )
)

app <- init(
  data = example_data,
  modules = modules(example_module())
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

The [teal.data::join_keys()] function is used to specify keys:

- [teal.data::join_keys()] is a collection of multiple [teal.data::join_key()] entries
- [teal.data::join_key()] specifies the relation between two datasets:
  - `dataset_1`, `dataset_2` - name of two datasets
  - `key` - (optionally) named vector of column names

Note that it is assumed that join keys are symmetric, i.e. `join_key("x", "y", "x_col" = "y_col")` will enable merge
from "x" to "y" and vice-versa.

For more information about preprocessing, reproducibility, relationships between datasets and DDL, please refer to the [`teal.data` package](https://insightsengineering.github.io/teal.data).
