---
title: "Teal and Bootstrap Themes"
author: "Coredev"
date: "2022-09-08"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Teal and Bootstrap Themes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

We offer the easy application of a custom bootstrap theme to a `teal::init` app. 
Teal is a consumer of the `bslib` R package which provides tools for customizing Bootstrap themes e.g. of Shiny apps.

## Usage

Teal users take the benefit of custom bootstrap themes by specifying the `teal.bs_theme` R option, which has to be set to `bslib::bs_theme` object.
The `bslib::bs_theme(...)` function creates a Bootstrap theme object, where you should specify the (major) Bootstrap version (default or one of 3, 4 or 5).
Optionally you could choose a **`bootswatch` theme** and **customize the app css** with functions like `bslib::bs_add_rules`.
Please read more about custom themes in [the `bslib` getting started vignette](https://rstudio.github.io/bslib/articles/bslib.html).
The `teal.bs_theme` R option has to be specified at the top of the code script.

Please install `bslib` package if not already installed

### `teal.bs_theme` R option

```
options("teal.bs_theme" = bslib::bs_theme("Custom Options"))

#######################
# teal::init() app code
#######################
```

### bootstrap version and themes

The best way to **explore** the bootstrap themes is to use `bslib::run_with_themer(shinyApp(app$ui, app$server))` or `bslib::bs_theme_preview()`,
both functions offer an interactive explore mode.
Please, remember that the interactive theming for Bootstrap 3 is not supported.
The `bslib::bs_theme_preview()` is recommended when the end user does not have any shiny app yet. 
When you already have a shiny app and want to test different bootstrap themes (and css styling) then `bslib::run_with_themer(shinyApp(app$ui, app$server))` is recommended.

Available bootstrap versions could be checked with `bslib::versions()` and bootstrap themes (`bootswatch`) with `bslib::bootswatch_themes()`.

```
# bslib::bootswatch_themes()
# bslib::versions()
options("teal.bs_theme" = bslib::bs_theme(version = "5", bootswatch = "lux")
# or
options("teal.bs_theme" = bslib::bs_theme_update(bslib::bs_theme(version = "5"), bootswatch = "lux"))
```

### custom teal css

The most important `teal` html tags have a proper id/class, so they could be directly styled.
The `bslib::bs_add_rules` function could be used around the `bslib::bs_theme` object to apply custom css rules.

```
library(magrittr)
options("teal.bs_theme" = bslib::bs_theme(version = "5") %>% bslib::bs_add_rules("Anything understood by sass::as_sass()"))
```

Other `bslib::bs_add_*` family functions could be used to specify low-level bootstrap elements.

### reset the bootstrap theme

Please use the `options("teal.bs_theme" = NULL)` call to return to the default shiny bootstrap for teal apps.

### theme not updated

One reason the theme is not updated could be the problem that a browser caches the previous one, especially when we run different themes one after another.
Please, use the "Cmd+Shift+r" (Mac) or "Ctrl+F5" (Windows) to hard refresh the webpage.

### regular `shiny::fluidPage`

If you want to update the theme in the regular `shiny::fluidPage` like app, you do not need the `teal.bs_theme` option. Then simply provide the `bslib::bs_theme` directly to the `shiny::fluidPage(theme = bslib::bs_theme(...), ...)`. 

### Examples

With `bslib::run_with_themer` to explore different themes.
Please, remember that the interactive theming for Bootstrap 3 isn't supported.


```{r, message=FALSE}
options("teal.bs_theme" = bslib::bs_theme(version = "5"))
library(teal)

app <- init(
  data = teal_data(
    dataset("IRIS", iris),
    dataset("MTCARS", mtcars)
  ),
  modules = modules(example_module(), example_module()),
  header = "My first teal application"
)

if (interactive()) {
  # Run with themer
  bslib::run_with_themer(shinyApp(app$ui, app$server))
}
```

Apply the already chosen theme with `options("teal.bs_theme" = bslib::bs_theme(...))`, in this example the `lux` theme from `bootstrap` version 5.

```{r, message=FALSE}
options("teal.bs_theme" = bslib::bs_theme(version = "5", bootswatch = "lux"))
library(teal)

app <- init(
  data = teal_data(
    dataset("IRIS", iris),
    dataset("MTCARS", mtcars)
  ),
  modules = modules(example_module(), example_module()),
  header = "My first teal application"
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

## Sources

- https://rstudio.github.io/bslib/
