---
title: "Teal as a Module"
author: "NEST CoreDev"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Teal as a Module}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

A Shiny developer interested in embedding Teal application into its own app, can use the Teal module composed of `ui_teal()` and `srv_teal()` functions. 
Unlike `init()`, this module will not automatically include session info footer, but it is possible to add it manually with `ui_session_info()` and `srv_session_info()`.
Using Teal as modules offers several advantages such as:

- Including one or multiple Teal applications in other app.
- Run Teal applications based on the dynamically created components like initial data, modules, filters.

# Example

Example below demonstrates how to embed Teal as a module in a Shiny application. 
Here, user can select dataset names which will be handed over and displayed in the Teal module. On the server side `srv_teal()` is called using the reactive `teal_data` object passed from the server of the parent app.

```{r setup, include=FALSE}
library(teal)
```
```{r app}
library(teal)

data <- teal_data() |> within({
  iris <- iris
  mtcars <- mtcars
  df <- data.frame(a = 1:10, b = letters[1:10])
})

mods <- modules(
  example_module("mod1"),
  example_module("mod2")
)

ui_app <- fluidPage(
  title = "Your app with teal as a module",
  selectInput("datasets", "Select datasets", choices = c("iris", "mtcars", "df"), selected = "iris", multiple = TRUE),
  ui_teal("teal", mods),
  ui_session_info("session_info")
)

srv_app <- function(input, output, session) {
  data_subset <- reactive(data[input$datasets])
  srv_teal("teal", data = data_subset, modules = mods)
  srv_session_info("session_info")
}

if (interactive()) {
  shinyApp(ui_app, srv_app)
}
```

```{r shinylive_iframe, echo = FALSE, out.width = '150%', out.extra = 'style = "position: relative; z-index:1"', eval = requireNamespace("roxy.shinylive", quietly = TRUE) && knitr::is_html_output() && identical(Sys.getenv("IN_PKGDOWN"), "true")}
code <- paste0(c(
  "interactive <- function() TRUE",
  knitr::knit_code$get("app")
), collapse = "\n")

url <- roxy.shinylive::create_shinylive_url(code)
knitr::include_url(url, height = "800px")
```
