---
title: "Actors"
author: "NEST CoreDev"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Actors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

There are two main actors in the `teal` development:

* `teal` App Developer
* `teal` Module Developer

## `teal` app developer

```{r, child="_setup.Rmd"}
```

```{r actors_mermaid1, echo=FALSE}
shiny::pre(
  class = "mermaid",
  "
%% This is a mermaid diagram, if you see this the plot failed to render. Sorry.
graph LR
    A[teal App Developer]--utilizes--> B[teal modules]
    B--to create--> C[teal app]
    D[app user]--uses--> C
style A fill:lightblue
style C fill:gold
style D fill:pink
"
)
```

The primary responsibility of a `teal` app developer is to leverage the available building blocks of the `teal` framework to create a functional `teal` app that analyses the data.

To expedite the app creation process, developers can take advantage of pre-existing `teal` modules found in `R` packages like `teal.modules.general` and `teal.modules.clinical`.

These modules are designed with a focus on standardization and versatility, making them suitable for a wide range of use cases.

When developing a `teal` app, the developer will select the most appropriate `teal` modules and integrate them into the app's interface to ensure seamless usability for end-users.

To learn more about the existing modules, visit [`teal.gallery`](https://insightsengineering.github.io/teal.gallery/), which contains several demo applications and their source code.

## `teal` module developer

```{r actors_mermaid2, echo=FALSE}
shiny::pre(
  class = "mermaid",
  "
%% This is a mermaid diagram, if you see this the plot failed to render. Sorry.
graph LR
    A[Teal App Developer]--utilizes--> B[teal modules]
    E[Teal Module Developer]--develops--> B
    B--to create--> C[teal app]
    D[app user]--uses--> C
style A fill:lightblue
style E fill:lightgreen
style D fill:pink
style C fill:gold
"
)
```

The main duty of a `teal` module developer is to construct a compatible `teal` module that can be utilized within the `teal` framework.

Several factors influence the scope and requirements for building a `teal` module.

When creating a reusable `teal` module, it's advisable to focus on making it as general and adaptable as feasible to maximize the possibilities of being re-used in the future.
However, developers have the freedom to create a `teal` module that is customized to suit the specific demands of a project.

Ultimately, one or more `teal` modules are employed to construct a `teal` app.

To learn more about creating custom modules follow the [Tutorial on Creating a Custom Module](creating-custom-modules.html).

## Workflow in a clinical trial study

```{r actors_mermaid3, echo=FALSE}
shiny::pre(
  class = "mermaid",
  "
%% This is a mermaid diagram, if you see this the plot failed to render. Sorry.
%%| fig-width: 7.5
graph LR
    subgraph Study B
        A2[Study Teal App Developer]--utilizes--> B2[teal modules]
        E2[Study Teal Module Developer]-.develops.-> B2
        B2--to create--> C2[study teal app]
        D2[study app user]--uses--> C2
    end
    E3[Teal Module Developer]--develops--> B
    E3--develops-->B2
    subgraph Study A
        A[Study Teal App Developer]--utilizes--> B[teal modules]
        E[Study Teal Module Developer]-.develops.-> B
        B--to create--> C[study teal app]
        D[study app user]--uses--> C
    end
style A fill:lightblue
style A2 fill:lightblue
style E fill:limegreen
style E2 fill:limegreen
style E3 fill:lightgreen
style D fill:pink
style D2 fill:pink
style C fill:gold
style C2 fill:gold
"
)
```

In a clinical trial study setting, a unique **study `teal` app developer** is assigned to each study team and is accountable for developing a tailored `teal` app for their respective study.

The **study `teal` app developer** will initially leverage existing `teal` modules from `R` packages created by **`teal` module developers**.

In cases where there is a need to create new modules tailored to the study, a **study `teal` module developer** will need to be involved.

Upon completion, each study team will have their own designated `teal` app tailored to their specific study.
