---
title: "Data Flow"
author: "NEST CoreDev"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Flow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, child="_setup.Rmd"}
```

```{r dataflow_mermaid1, echo=FALSE}
shiny::pre(
  class = "mermaid",
  "
%% This is a mermaid diagram, if you see this the plot failed to render. Sorry.
sequenceDiagram;
    autonumber
    participant data
    participant filters
    participant filtered data
    participant teal module
    data->filters: gets
    filters->>filtered data: becomes
    filtered data->>teal module: sent to
"
)
```

The sequence diagram above illustrates the different stages that data goes through within the `teal` framework, supported by the `teal.slice` R package:

1. Data is loaded into the `teal` app;
    - See the [Filter panel vignette](filterpanel.qmd) for details;
2. Data is passed to the filter panel;
    - Users _(or app developers)_ can specify filters to apply;
    - Filters can be specified globally, for the whole app, or for specific modules;
3. Filtered data is sent to `teal` modules for analysis;
    - Each modulev creates a [`qenv`](qenv.qmd) object track the code that is be applied to the data _(and used for reproducibility)_;

Whenever filters are added or removed, these stages are restarted, providing the `teal` module with new filtered data to conduct the required analysis.
